### 2018-07-07 paper/paper.tex
paper progress

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

%\documentclass[twoside,twocolumn]{article}
%\documentclass{article}[11pt]
\documentclass[fleqn]{article}

\usepackage{listings}

\lstset{
  xleftmargin=0em,
}
\usepackage{xcolor} %custom colours
\usepackage{mdframed} %nice frames

\definecolor{light-gray}{gray}{0.95} %the shade of grey that stack exchange uses

%Put in the main document

\usepackage{blindtext} % Package to generate dummy text throughout this template

\usepackage[sc]{mathpazo} % Use the Palatino font
\usepackage[T1]{fontenc} % Use 8-bit encoding that has 256 glyphs
\linespread{1.05} % Line spacing - Palatino needs more space between lines
\usepackage{microtype} % Slightly tweak font spacing for aesthetics

\usepackage{amsfonts}

\usepackage{amsmath}

\usepackage[english]{babel} % Language hyphenation and typographical rules

\usepackage[hmarginratio=1:1,top=32mm,columnsep=20pt]{geometry} % Document margins
\usepackage[hang, small,labelfont=bf,up,textfont=it,up]{caption} % Custom captions under/above floats in tables or figures
\usepackage{booktabs} % Horizontal rules in tables

\usepackage{lettrine} % The lettrine is the first enlarged letter at the beginning of the text

\usepackage{enumitem} % Customized lists
\setlist[itemize]{noitemsep} % Make itemize lists more compact

\usepackage{abstract} % Allows abstract customization
\renewcommand{\abstractnamefont}{\normalfont\bfseries} % Set the "Abstract" text to bold
\renewcommand{\abstracttextfont}{\normalfont\small\itshape} % Set the abstract itself to small italic text

\usepackage{titlesec} % Allows customization of titles
\renewcommand\thesection{\Roman{section}} % Roman numerals for the sections
\renewcommand\thesubsection{\roman{subsection}} % roman numerals for subsections
\titleformat{\section}[block]{\large\scshape\centering}{\thesection.}{1em}{} % Change the look of the section titles
\titleformat{\subsection}[block]{\fontsize{10}{10}\bfseries}{\thesubsection.}{1em}{} % Change the look of the section titles

\usepackage{fancyhdr} % Headers and footers
\pagestyle{fancy} % All pages have headers and footers
\fancyhead{} % Blank out the default header
\fancyfoot{} % Blank out the default footer
\fancyhead[C]{Controlled Natural Language and Robotics $\bullet$ Calvin Claus} % Custom header text
\fancyfoot[RO,LE]{\thepage} % Custom footer text

\usepackage{titling} % Customizing the title section

\usepackage{hyperref} % For hyperlinks in the PDF

\lstset{basicstyle=\small}

%----------------------------------------------------------------------------------------
%	TITLE SECTION
%----------------------------------------------------------------------------------------

\setlength{\droptitle}{-4\baselineskip} % Move the title up

\pretitle{\begin{center}\Huge\bfseries} % Article title formatting
  \posttitle{\end{center}} % Article title closing formatting
\title{Controlled Natural Languages and Robotics} % Article title
\author{%
  \textsc{Calvin Claus (1429713)}
  \normalsize TU Wien\\ % Your institution
  \normalsize \href{mailto:e1429713@student.tuwien.ac.at}{e1429713@student.tuwien.ac.at} % Your email address
}
\date{\today} % Leave empty to omit a date
\renewcommand{\maketitlehookd}{%
  \begin{abstract}
    \noindent This paper gives a brief introduction into Controlled Natural Languages (CNLs), providing an overview over common language-rules and outlining the long term goals of CNL research. It describes how to construct a simple CNL interface to allow robots to receive human commands. Finally, it analyzes if a robot can successfully build a useful knowledge base over time, using CNL input by a human speaking naturally. It finds that, at the current state of CNL research, CNL can be used to build a good knowledge base, but the human has to unnaturally adapt his/her speaking to the fact that a robot should understand him/her, even though he/she is conforming to CNL rules.
  \end{abstract}
}

%----------------------------------------------------------------------------------------

\begin{document}

% Print the title
\maketitle

%----------------------------------------------------------------------------------------
%	ARTICLE CONTENTS
%----------------------------------------------------------------------------------------

\section{Introduction}

\lettrine[nindent=0em,lines=3]{I} t's a fascinating idea:

\section{Ivy}
Ivy is a system that is designed to verify safety of infinite-sate systems.
It aims to assist protocol architects, algorithm authors, and the like in proofing the safety and correctness of their algorithm or protocol.

\subsection{How Ivy is used}

To understand how Ivy can be used, a common workflow that incorporates Ivy in algorithm verification is presented:

\begin{enumerate}
  \item design the algorithm
    \begin{itemize}
      \item purely creative task
      \item Ivy offers no support in this area
    \end{itemize}
  \item encode it in Ivy language
    \begin{itemize}
      \item a one to one translation is most likely not possible
      \item user must be ready for the paradigm and reasoning change when writing Ivy language as opposed to a procedural programming language
    \end{itemize}
  \item verify the properties one designed the algorithm to fulfil
    \begin{itemize}
      \item specify properties as invariants
      \item make invariants inductive
    \end{itemize}
\end{enumerate}

\subsubsection{Step 1 - algorithm design}
Step 1 is executed entirely independently of Ivy. The Ivy language or the Ivy system do not provide any support in actually coming up with an algorithmic solution to a problem.
The step is mentioned here to show the context the Ivy system can be used in: to answer the question "is an algorithm correct"?

\subsubsection{Step 2 - encoding in Ivy language}
Step 2 requires the user to understand the "unusual" \cite{refLanguageDoc} - compared to a procedural programming language -  Ivy language features. The user must
find a way to use the Ivy language features to encode not only the algorithm itself but also the "system context" or "framework" in which the algorithm was designed to execute.
For example, to reason about distributed consensus algorithms Lynch has defined a "context" or "framework" in which distributed algorithms operate.
This is explained in detail later on in this work. Part of this framework describes that the network is defined as a graph, where edges can hold messages.
The system executes two distinct steps: one in which messages can be sent and another in which the states of nodes are altered.
The point is, both the algorithm and the framework in which it operates have to be flawlessly described through the Ivy language in this step. This process is
fully creative and will - judging by the authors experience - most likely not allow a mindless one-to-one translation from a procedural algorithm.
In this step the user also has to tell Ivy which methods (or "actions" in Ivy speak), e.g. \textit{connect\_to\_server(node)}, the environment can execute, . These actions are called "exported".
These are the actions the final user of the algorithm would call, and it is our desire to proof correctness of these methods. Furthermore the user can supply
the program with assertions, that they wish to proof cannot fail.

\subsubsection{Step 3 - verification}
The verification happens in step 3. The goal is to specify inductive invariants that describe the properties to be proven about the algorithm. "Inductive" is an important
keyword here. An invariant is inductive iff the following three properties are satisfied:
\begin{enumerate}
  \item initiation: the invariant is true in all legal initial states
  \item safety: for any program state that satisfies the invariant - regardless of whether this state is reachable - no exported action can cause an assertion failure starting in that state
  \item consecution: for any program state that satisfies the invariant - regardless of whether this state is reachable - then after executing any exported action, the invariant remains satisfied
\end{enumerate}
Inductive invariants is one of the more difficult concepts to get ones head around, especially as a user who does not have any experience with automatic verification.
Ivy does not consider the reachability of a system state.  As long as the state is in compliance with the invariants it is assumed to be reachable.
The author of this paper needed some time to understand why Ivy generates counterexamples that are clearly not reachable, e.g. negative values for counters that are set to zero at initialization.
To fix this one would have to make such the counter part of the invariant e.g. $invariant counter >= 0$.

The name "Ivy" stems from "interactively verifying", which highlights the key differentiator of Ivy to Coq or Dafny.
Ivy actively supports the user in the search for an inductive invariant by graphically displaying a concrete minimal counterexample to induction.
The user can then generalize from the counterexample to strengthen the invariants. This counterexample $\rightarrow$ strengthened invariant $\rightarrow$ counterexample process continues
until an inductive invariant is found and the correctness of the algorithm is proven. However, there is no mechanism in place that automatically disallows invariants that exclude reachable states.
This means that it's easy to exclude theoretically reachable edge cases one hasn't thought of, by writing invariants that are too strong.

\subsection{The Ivy language}
The Ivy language is a core part of the Ivy system. It is the language through which the algorithm to be verified is defined.
This section presents a selection of core language features that should help the reader understand the paradigm of modelling in Ivy, as well as lay the necessary foundation to understand
the rest of the work presented in this paper. This section oversimplifies and does not give an exhaustive documentation, which is given in \cite{refLanguageDoc}.

\subsubsection{Types and declarations}
In Ivy one mainly uses uninterpreted types. ``Uninterpreted'' means that nodes in the system state are an arbitrary set with at least one element. The cardinality of this
set is parameterized.  Meaning how many nodes are generated to construct a counter example to induction is entirely up to Ivy, one does not manually instantiate instances of these types.

\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
\begin{lstlisting}
type node
\end{lstlisting}
\end{mdframed}

\noindent To relate instances of types to each other we can declare functions.

\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
\begin{lstlisting}
type node
type value
function value(N:node) -> value
\end{lstlisting}
\end{mdframed}

\noindent A function that maps to bool can be declared as a relation, so

\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
\begin{lstlisting}
function message_received(RECEIVER:node, SENDER:node) -> bool
\end{lstlisting}
\end{mdframed}

\noindent is the same as

\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
\begin{lstlisting}
relation message_received(RECEIVER:node, SENDER:node)
\end{lstlisting}
\end{mdframed}

\subsubsection{Assignments}

To assign values to relations, functions or variables $:=$ is used. To assign to multiple values
at the same time one can use an uppercase placeholder. For example setting all nodes with an incoming message to failed:

\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
\begin{lstlisting}
failed(X) := ~incoming_message(X);
\end{lstlisting}
\end{mdframed}

\subsubsection{Actions}
To define a callable action, that is callable from within another action or from the environment if it is exported
this is the syntax.

\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
\begin{lstlisting}

action fail_node_and_send(x:node) = {
  fail(x) := true;
  call send;
}
export fail_node_and_send

action send = {
  sent(X) := true;
}
\end{lstlisting}
\end{mdframed}

\subsubsection{Non deterministic choice}
To assign to a variable, relation or function a value non-deterministically (randomly) one uses ``*''.
This line non-deterministically chooses a number of nodes to have failed.

\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
\begin{lstlisting}
failed(X) := *;
\end{lstlisting}
\end{mdframed}

\subsubsection{Assume, require and ensure}
To specify safety conditions in an action we can use require and ensure.
They both act like a traditional assert in that it does not allow control to pass if it is false.
They are different semantically in that require assigns blame to the actions caller, while ensure
is a condition that the action itself must guarantee. This is especially relevant to restrict
the environment in the way it can call an action.

Assume is different in that it does not allow control to pass through if the condition is not satisfied.
The author has found this to be mostly useful in combination with non-deterministic assignments that one
wants to restrict.

The following action chooses a random node to assign to y, but will not allow control to pass if
the chose node has failed. Therefore the actions will choose a random non failing node to send a message to x.
If there is no node that hasn't failed, this action cannot terminate.

\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
\begin{lstlisting}
action random_node_sends_message_to(x:node) = {
  y := *;
  assume ~failed(y);
  send(x, y);
}
\end{lstlisting}
\end{mdframed}

Assume simplifies modelling in some situations, as it can easily eliminate impossible behaviour. However,
it is risky to use assume for the same reason: one might dismiss a valid counterexample to the specification
by assuming it cannot happen.

\subsubsection{Invariant}
To specify the properties we want to proof about our algorithm we encode them in inductive invariants like so:

\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
\begin{lstlisting}
invariant system.rounds >= 0
\end{lstlisting}
\end{mdframed}

These must hold at all times when an action is not executing, thus they have to be initially true and each of
the exported actions must preserve them. However, an invariant is not guaranteed to hold when an action calls another action.

\subsection{Decidability}
Not all first-order formulas are decidable in Ivy. Therefore, the system defines a subset of first-order formulas as its decidable fragment.
This not only constraints the structure of logical formuals but also functions. In Ivy only stratified functions are allowed. Meaning that if a graph was constructed
with every type being a node and every function from a type A mapping to another type B would be represented as an edge the graph has to be cycle free. Thus there cannot be a function from
$X \rightarrow Y$ and from $Y \rightarrow X$ in the same specification.

Ivy is extremely vocal about this and will let the user know if his specification does not lie in the decidable fragment.

\section{Distributed Consensus Algorithms with stopping failures}

The algorithm that was implemented in Ivy as part of this work is called \textit{Floodset} and was taken out of Lynch's work ``Distributed Algorithms'' \cite{refNancy}.
\textit{Floodset} is a consensus algorithm for synchronous networks that is designed to reach consensus in a network with failing processes.
To understand \textit{Floodset}, one needs to first be made familiar with the aforementioned definition of the framework/system within which this algorithm operates

\subsection{Framework - Synchronous Network Systems}

The network of processes is represented as a directed graph $G=(V,E)$. Where $n$ is $|V|$ the number of nodes in the network. Each node represents a process in the network.
The communication paths between nodes are represented by directed edges in the graph.
$out-nbrs_i$ is the set of nodes to which an edge from node $i$ exists.
$in-nbrs_i$ is the set of nodes from which an edge to node $i$ exists.
$M$ denotes the message alphabet, meaning that elements of $M$ can be used to construct messages (vectors of elements of $M$) from nodes to other nodes.
There are a number of components associated with each node:
\begin{itemize}
    \item $states_i$, a set of states (can be infinite)
    \item $start_i$, a nonempty subset of $states_i$
    \item $msgs_i$, a message generation function that maps $states_i X out-nbrs_i$ to elements of $M U {null}$
    \item $trans_i$, a state transition function that maps $states_i X vectors of M index by in-nbrs_i$ to $states_i$
\end{itemize}
Thus, a node is always in some state in $states_i$. Importantly, the set of states is not necessarily finite, meaning that unbounded data structures like arrays can be modeled in the system.
Each node defines a message generation function that uses the current state to return a message for each out neighbour. Similarly each node has a state transition function that is used to determine
from the current state and all incoming messages the state the next state the node will transition to.

The edges in the graph are called channels and can hold at most one message at a time or $null$ to denote an empty channel.
The system of nodes operates in steps. It begins with all nodes in an arbitrarily chosen start state from $start_i$ and all channels empty.
The processes then all repeatedly perform the following two steps \textit{synchronously}:
\begin{itemize}
  \item Apply the message generation function inputting the current state and one of the outgoing neighbours. Put these message in the channel to the chosen outgoing neighbour. Do this for each outgoing neighbour.
  \item Apply the state transition function to the current state and the array of incoming messages. Transition the node to the returned state. Remove all messages from the channels.
\end{itemize}
\textit{Synchronously} in this case means that all nodes wait for each other to finish step1 before executing step2 and vice versa.
The combinations of these two steps is a \textit{round}.

\subsection{Stopping failures}
The system presented in Lynch's work \cite{refNancy} is capable of modelling failure in a variety of ways.
This work makes use of stopping failure only. This kind of failure occurs when a process stops in the middle of its execution. With regards to the presented framework this
is modeled as a process failing before or after performing some instance of Step 1 or Step 2. Additionally, a process may also stop in the middle of performing Step 1. This allows
the interesting situation in which a process has communicated its messages to some but not yet all processes. It is important to note that processes are not thought of performing
the message generation operation sequentially. Instead, we say that the order in which messages are put in the channels is non-deterministic. This clarification, allows a failing process
to send its messages to an arbitrary subset of its outgoing neighbours only.

\subsection{Floodset}
The problem at hand is as follows: An arbitrary number of processes start with individual inputs from a value set $V$. This is modeled by each node having at least one start state
for every possible value $v ∈ V$. Also, a default value $v_0 ∈ V$ that is equal for all nodes is defined.
Processes communicate with each other through several \textit{rounds}. The nature of their communication and its effect on their states is defined by their $msgs_i$ and $trans_i$ functions.
Nodes may fail according to the \textit{stopping failure} model described above. The total number of failing processes is bounded by a constant $f$. The entire network knows the value of $f$ and
can therefore utilise it in their state transition function.
All processes that did not fail must eventually decide on a single value from the value set $V$. \textit{Deciding} is modeled by nodes setting a special \textit{decision} state component to a value in $V$.
Their decision is subject to the following correctness conditions:
\begin{itemize}
  \item \textbf{Agreement}: No two processes decide on different values.
  \item \textbf{Validity}: If all processes start with the same initial value $v ∈ V$, then $v$ is the only possible decision value
  \item \textbf{Termination}: All nonfaulty processes eventually decide.
\end{itemize}
We assume the processes are in a complete, undirected graph. This means that every node has complete knowledge over the network. Channels are fully reliable - every message gets delivered.

In the \textit{Floodset} algorithm processes hold a set $W$ as one of their state components. This set holds the subset of values of $V$ that the process has seen. Every node's initial $W$ holds only its initial value.
In Step 1 of every round nodes broadcast their $W$. In Step 2 all $v ∈ V$ that \textit{(a)} are sent to a process and \textit{(b)} the process does not yet hold in its $W$ are added to its $W$. In the $f+1$th round
a node decides: If $W$ is a singleton set then it decides on the unique element of $W$, otherwise it decides on the default value $v_0$.

\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
\noindent $states_i$:
\begin{gather*}
  rounds   \in  \mathbb{N}, \ \mathrm{initially} \ 0 \\
  decision  \in  V \cup \{ unknown \}, \ \mathrm{initially} \ unknown \\
  W \subseteq V, \ \mathrm{initially \ the\  singleton\  set\  consisting\  of\  i's\  initial\  value}
\end{gather*}
\end{mdframed}

\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
\noindent $msgs_i$:
\begin{gather*}
  if\ rounds \le f\ \mathrm{send}\ W\ \mathrm{to\ all\ other\ processes}
\end{gather*}
\end{mdframed}

\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
\noindent $trans_i$:
\begin{gather*}
  rounds\ :=\ rounds+1\\
  \mathrm{let}\ X_j\ \mathrm{be\ the\ message\ from\ j,\ for\ each\ j\ from\ which\ a\ message\ arrives}\\
  W\ :=\ W \cup \bigcup_{j} X_j\\
  \mathrm{if}\ rounds = f+1\ \mathrm{then}\\
  \ \ \   \mathrm{if}\ |W| = 0\ \mathrm{then}\ decision := v,\ \mathrm{where}\ W = {v}\\
  \ \ \   \mathrm{else}\ decision := v_0
\end{gather*}
\end{mdframed}

\noindent What follows is a proof of the correctness of the \textit{Floodset} algorithm to illustrate the principles behind it to the reader.

Let $W_i(j)$ denote the value of $W$ for node $i$ in round $j$. The following Lemmas apply:\\

\noindent \textbf{Lemma 1}:\\
If no process fails during a particular round $r,\ 1 \le r \le f + 1$, then $W_i(r) = W_j(r)$ for all $i$ and $j$ that did not fail after $r$ rounds.\\
\textbf{Proof}:\\
\textit{Indirect}: If there exists a value $v \in V$ that is in $W_i(r)$ but not in $W_j(r)$ after round $r$ in which no process failed, that implies that some node $k$ sent a value to $i$ that it did not send to $j$, or $i$ failed to send a member of $W_i(r-1)$. Since all nodes broadcast their $W$, this is only possible if node $k$ or $i$ failed to send at least one message. But since no processes failed, each process must have shared their $W$ with every other process. \textit{Contradiction}. \\

\noindent \textbf{Lemma 2}:\\
Suppose that $W_i(r) = W_j(r)$ for all $i$ and $j$ that did not fail after $r$ rounds. Then for any round $r',\ r \le r' ≤ f + 1$, the same holds, that is, $W_i(r’) = W_j(r’)$ for all $i$ and $j$ that did not fail after $r’$ rounds.\\
\textbf{Proof}:\\
Since all active nodes  have equal $W$s after $r$ rounds, there exists no $v \in V$ that is in some $W_i(r),\ i\ \mathrm{some\ active\ node}$ that is not in every other $W_j(r),\ j\ \mathrm{any\ active\ node}$.
Since nodes will only send their $W$s, this implies that no node can send a message in some future round $r'$ that contains a value $v$ that some node does not have in its $W$, thus $W_i(r’) = W_j(r’)$, $i,\ j$ not failed.\\

\noindent \textbf{Lemma 3}:\\
After $f+1$ rounds there has been a round $r'$ in which no process failed\\
\textbf{Proof}:\\
\textit{Indirect:} If a process failed in all of $f+1$ rounds at least on process must have failed in all of the $f+1$ rounds. This implies
that at least $f+1$ processes failed. Only $f$ process can fail. \textit{Contradiction}.\\

\noindent \textbf{Lemma 4}:\\
If processes $i$ and $j$ are both active after $f+1$ rounds, then $W_i = W_j$ at the end of round $f + 1$.\\
\textbf{Proof}:\\
It follows from \textit{Lemma 3}, that after $f+1$ rounds at least one round, $r'$, in which no process failed, occurred. From \textit{Lemma 1} we therefore derive that $W_i(r') = W_j(r')$ for all active $i,\ j$.
From \textit{Lemma 2} it follows that $W_i(f+1) = W_j(f+1)$. \textit{Q.e.d}.\\

\noindent The Lemmas allow proofing the theorem:\\
\noindent \textbf{Thorem}:\\
\textit{Floodset} satisfies the correctness conditions.\\
\textbf{Agreement}:\\
From \textit{Lemma 4} we know what $W_i = W_j$ after $f+1$ rounds. Since all decision conditions are the same and only depend on $W$, agreement is satisfied.\\
\textbf{Validity}:\\
If all processes start with the same initial value $v \in V$, then there exists no value $v', v \neq v'$, that is part of any $W$. Thus, from $W_i(f+1) = W_j(f+1)$, $i,\ j$ active, and $W_i(f+1) = \{v\}$ it follows
that every active node decides $v$.\\
\textbf{Termination}:\\
Trivially, nodes decide after $f+1$ rounds.\\

This work now presents a way to translate Lynch's framework and the \textit{Floodset} algorithm to a specification in the Ivy language.

\section{Specification}
To summarize: at hand we have a parameterized system with $n$ process and $f$ number of faults. We want to show \textit{Floodset} is correct regardless of the parameters. It is trivial to verify for a small number of $n,\ f$, but the question is ``does \textit{Floodset} work for any assignment of $n,\ f,\ n < f$''.
To answer this question we are using the theorem proofer Ivy; first encoding the algorithm in the Ivy language and then constructing inductive invariants to proof the correctness.

The goal of this section is to walk through the specification step by step, to give the reader an understanding of the paradigms relevant to automatic verification and Ivy in particular. The complete code can be found here \cite{github}.

\subsection{Modeling strategy}

\subsubsection{Modularity}
When verifying an algorithm like \textit{Floodset} in Ivy, an algorithm designer would most likely prefer to be able to translate their original algorithm to the Ivy language in as close to a one-to-one fashion as possible. However, this is not quite convenient for \textit{Floodset} and the framework it operates in.
To separate system and nodes one would have to rely heavily on iterations. E.g. for \textit{Step 1} the system would iterate over all nodes and let each node take care of sending it's own message.

\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
\begin{lstlisting}
action round = {
	var iter := node.t.iter.create(0);
  while ~iter.is_end
	{
 	 call send_messages(iter.val)
	 iter := iter.next;
	}
}
action send_messages(n:node.t) = {
  node.message(n, V) := node.w(n, V);
  node.incomming_message(O, n) := true;
}

export round
\end{lstlisting}
\end{mdframed}

However, the code above does not actually execute an iteration over all nodes as one would expect. Looking at an Ivy trace we see what really happens:

\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
\begin{lstlisting}
message := *;
incomming_message := *;
\end{lstlisting}
\end{mdframed}

Ivy non-deterministically assigns to all relations involved in the loop, instead of actually executing the loop.  This is because proofing correctness of loops can only be done through inductive loop invariants specified by a human. Ivy therefore allows specifying invariants together with a loop.
\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
\begin{lstlisting}
  while ~iter.is_end
  invariant node.message(N,V) = node.w(N,V)
  invariant N<=iter.val -> incomming_message(O, N)
	{
    . . .
	}
\end{lstlisting}
\end{mdframed}

These invariants must hold before the loop and after every iteration. The following is a trace of the execution of the loop with the specified invariants. First they are asserted, then the relations are set to non-deterministic values and finally an assume is used to restrict control to values for the relation that fulfil the invariants.

\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
\begin{lstlisting}
assert message(N,V) = w(N,V)
assert N:t <= t.iter.val(loc:iter) -> incomming_message(O,N)
incomming_message := *
message := *
assume message(N,V) = w(N,V)
assume N:t <= t.iter.val(loc:iter) -> incomming_message(O,N)
\end{lstlisting}
\end{mdframed}

As a result of this architecture the program logic has to be specified twice: once in the form of invariants and another time in the actual loop. Such a program therefore has two possible origins for any error: a fault in the algorithm or an inadequate invariant. This structure also increases the difficulty of debugging, because Ivy only shows unhelpful $assert$, $assume$ and  non-determinsitic assignments ($:= *$) in its traces, rather than execution of the actual lines in the $send_messages$ function.

For the simple \textit{Step 1} this approach may still be considered viable, but for the more complex \textit{Step 2} and the failure simulation logic, the author has found that the duplication of logic and the decreased debugging efficiency makes the increased modularity achieved through iterations not worth the mental and maintenance effort that the loop invariants bring with them.

For this reason the following specification of \textit{Floodset} uses a non-modular structure where single lines predominately alter the state of the system as a whole.

\subsubsection{Cardinality relationships}
There is no cardinality operator in IVy. In set notation we would write something like $|\{n|node(n) \land failed(n)\}|$ to count the number of failed nodes in a system. When programming in a standard procedural language we would use a simple loop for this task but, as has been explained in the previous section, loops are somewhat problematic in Ivy. The author couldn't find a way to specify an inductive loop invariant for a loop counting the number of failed nodes without a cardinality operator at hand. This means that alternative encodings for the assumption $f<n$ and the statement $if rounds = f+1$ must be found.

The assumption $f<n$, ``number of failed nodes is bounded to an amount smaller than the number of total nodes'', can alternatively be expressed via $\exists n\ \neg failed(N)$.

To figure out how to alternatively encode $if rounds = f+1$ deep understanding of the algorithm is necessary. One must realise that $rounds = f+1$ entails that a ``clean round'' occurred (see \textit{Lemma 3}). Where ``clean round'' describes ``a round in which no node failed''. The processes are ready to decide after a clean round has occurred. They can decide either immediately, or an arbitrary number of rounds after the clean round. This can be modeled with non-determinism:

\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
\begin{lstlisting}
clean_round_occured := (clean_round_occured | (forall N . ~node.failed_during_round(N)));
if clean_round_occured {
  decision_round := *;
};
\end{lstlisting}
\end{mdframed}

And an invariant that describes the properties of a clean round, that contains the premise of \textit{Lemma 1}:

\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
\begin{lstlisting}
# lemma 1:
invariant system.clean_round_occured -> ((~node.failed(A) & ~node.failed(B)) -> (node.w(A, V) = node.w(B, V)))
\end{lstlisting}
\end{mdframed}

The (other) chosen invariants will be explained in detail later on in this paper.

\subsection{Types, functions, relations}
First the types, functions, relations and individual variables that are used throughout the program are presented:

\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
\begin{lstlisting}
type node
type value
object node = {
  #graph/system related
  function val(X:node) : value
  function default(X: node) : value
  individual def : value
  relation failed(A:node)
  relation failed_before_round(A:node)
  relation failed_during_round(A:node)
  relation has_single_value(N:node)

  #state
  relation w(SELF:node, V:value)
  relation decision(SELF:node, VALUE:value)

  #communication
  relation message(N:node, V:value)
  relation incomming_message(SELF:node, OTHER:node)
  relation incomming_message_failing(SELF:node, OTHER:node)
}

object system = {
  individual decision_round:bool
  individual clean_round_occured:bool
}
\end{lstlisting}
\end{mdframed}

Their meaning will become clear as the code is presented in the following paragraphs.

A round in the specification is defined as follows:
\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
\begin{lstlisting}
action round = {
  require ~node.decision(N,V) & ~decision_round;
  node.failed_before_round(N) := node.failed(N);
  call step1;
  call simulate_failure;
  clean_round_occured := (clean_round_occured | (forall N . ~node.failed_during_round(N)));
  if clean_round_occured {
    decision_round := *;
  };
  call step2;
}
export round;
\end{lstlisting}
\end{mdframed}

Line 2 makes it illegal for the environment to call \textit{round} if decisions already occurred.
Line 3 does some bookkeeping.
\textit{call step1} models the communication step. \textit{simulate\_failure} simulates failure before, during
or after \textit{step1}. After \textit{decision\_round} has been set as explained above, \textit{call step2} is
executed to model the state transition of the nodes.

\subsection{Step 1 - Message sending}
The contents of a nodes message is modeled via the relation $message(N:node, V:value)$. Where
$message(1, 0) = true; message(1, 1) = true;$ means node $1$ includes in its message the values $0$ and $1$.
Since a node never sends the same message to all nodes for any given round we don't need to include the
recipient in the relations.

The actual delivery of a message is modeled via the $incomming_message(RECIPIENT:node, SENDER:node)$ relation. Where $incoming_message(0, 1) = true; incoming_message(0, 2) = true;$ means that node $0$ receives the messages from node $1$ and node $2$.

These relations allow us to use Ivy's assignments to easily model \textit{Step 1} in a round of \textit{Floodset}.
Line 2 assigns for all nodes $A$ and values $V$ to $message(A,V)$ the boolean value of $node.w(A,V) \land \neg failed(A)$. In other words every node that hasn't failed prepares a message with all values in their $w$. Line 3 states that all nodes that haven't failed send messages to all other nodes.

\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=15, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
\begin{lstlisting}
action step1 = {
  node.message(A, V) := node.w(A, V) & ~node.failed(A);
  node.incomming_message(O, A) := ~node.failed(A);
}
\end{lstlisting}
\end{mdframed}

\subsection{Step 2 - Message sending}

Line 2 of the \textit{step2} action adds to $w$ of node $N$ all values $V$ for which a node $J$ exists that has sent a message to $N$ that contains $J$.
The conditional statement in Line 4 corresponds to $if\ rounds = f+1$ in the original algorithm. Nodes with a single value in $w$ decide on that value, otherwise
they decide on the default value.
Finally, for good measure, $incomming\_message$ and $message$ are reset, but this is not necessary for the proof.

\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=10, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
\begin{lstlisting}
action step2 = {
  node.w(N,V) :=
    node.w(N,V) | (exists J . node.incomming_message(N,J) &
                              node.message(J,V));

  if decision_round {

    node.has_single_value(N) :=
      (exists J . node.w(N,J)) &
      (forall V,O . O=V | (node.w(N,V) -> ~node.w(N,O)));

    node.decision(N,K) :=
      node.has_single_value(N) & node.w(N,K) & ~node.failed(N);

    node.decision(N,node.default(N)) :=
      (~node.has_single_value(N) | node.decision(N,node.default(N)))
      & ~node.failed(N);

  };

  node.incomming_message(A,B) := false;
  node.message(A,V) := false;
}
\end{lstlisting}
\end{mdframed}

\subsection{Failure simulation}

The \textit{simulate\_failure} action simulates the failure model described by Lynch.
It is called after \textit{step1} and illustrates sophisticated usage of the $assume$ functionality in Ivy.
In line 2 it choses a random number of nodes to fail. The assume in line 3 however excludes all
cases where a node that has already failed was chosen to fail in this round.
Line 4 prepares for each node $M$ a non-deterministically chosen set of nodes that it sends a message to.
Line 6 updates the \textit{incomming\_message} relation. If node $M$ failed during this round, the set of nodes
 $M$ sends messages to is changed to the values of \textit{incomming\_message\_failing}, otherwise it remains the same.
 After the nodes that \textit{failed\_during\_round} are set to \textit{failed} in line 12, line 13 assumes
 the condition $f<n$. This excludes the cases where all nodes failed.

\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=5, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
\begin{lstlisting}
  action simulate_failure = {
    node.failed_during_round(N) := *;
    assume node.failed_during_round(N) -> ~node.failed(N);
    node.incomming_message_failing(N,M) := *;

    node.incomming_message(N,M) :=
      (node.failed_during_round(M) &
        node.incomming_message_failing(N,M)) |
      (~node.failed_during_round(M) &
        node.incomming_message(N,M));

    node.failed(N) := (node.failed(N) | node.failed_during_round(N));
    assume exists N . ~node.failed(N);
  }
\end{lstlisting}
\end{mdframed}

Finally, we describe the initial system state in a special \textit{after init} action. When
proofing correctness Ivy requires the initialisation to entail the invariants. All relations and functions
that are not assigned any values here are set automatically by the environment. E.g. $node.val(N)$ is
treated as a parameter the environment sets.

\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=5, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
\begin{lstlisting}
after init {
  def := *;
  default(N) := def;
  w(N, V) := V=node.val(N);
  incomming_message(A, B) := false;
  decision(N, V) := false;
  clean_round_occured := false;
  decision_round := false;
  node.failed_before_round(N) := false;
  node.failed(N) := false;
  node.has_single_value(N) := false;
}
\end{lstlisting}
\end{mdframed}

\subsection{Invariants}

Now that the system has been modeled the next step of verifying correctness with Ivy is specifying inductive invariants. The simplest invariants to come up with are those that describe the correctness of the algorithm at hand. In our case these are \textit{agreement}, \textit{validity}, \textit{termination}.

\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=5, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
\begin{lstlisting}
####### Correctness Conditions #######

# Agreement: No two processes decide on different values.
invariant (~node.failed(A) & ~node.failed(B)) ->
            (node.decision(A,V) = node.decision(B,V))

# every node decides on at most one value
invariant node.decision(N,V) -> (~exists O . node.decision(N,O) & O~=V)

# Validity: If all processes start with the same initial value v,
#           then v is the only possible decision value
invariant (exists N,M . node.val(N) ~= node.val(M)) |
          (system.decision_round ->
            (~node.failed(N2) -> node.decision(N2,node.val(N2))))

# Termination: All nonfaulty processes eventually decide.
invariant system.decision_round <->
            (forall N . (~node.failed(N) -> (exists V . node.decision(N,V))))
\end{lstlisting}
\end{mdframed}

If one were to run $ivy_check$ on the program without specifying any
additional invariants Ivy finds that the \textit{agreement} invariant is not inductive
and produces the a counter example to induction (CTI). One can display the CTI in a GUI.
In the left hand section of the program we can see that the system transition from state 0
to state 1 by executing \textit{round}. In the center Ivy displays state 0, meaning the state
before \textit{round} was executed. The brown edges specify the relation $w$, the purple edges
show $val$.

---SCREENSHOT 1 TODO---

By examining state 0 we realise that no node has their initial value modeled by $val$
in its $w$. The programmer must realise this is an unreachable state and exclude it by coming
up with an inductive invariant. The process of finding invariants interactively is supported by Ivy but
beyond the scope of this paper.

An invariant excluding the unreachable state is one that states that the initial value of a node is always in its $w$:
\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=5, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
\begin{lstlisting}
# initial value is in W
invariant node.w(N, node.val(N))
\end{lstlisting}
\end{mdframed}

The method for finding the remaining invariants is identical:
\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=5, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
\begin{lstlisting}
# the global default is the one the nodes know
invariant node.default(N) = node.def
# if a value is in a process' W, it must be initial value of some other process ( no invented values )
invariant node.w(N, V) -> (exists M . node.val(M)=V)
# lemma 1:
invariant system.clean_round_occured -> ((~node.failed(A) & ~node.failed(B)) -> (node.w(A, V) = node.w(B, V)))
\end{lstlisting}
\end{mdframed}

Since it is also required that the framework behaves correctly it makes sense to include some
correctness conditions specifically aimed at the system.

\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=5, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
\begin{lstlisting}
####### system correctness checks #######

# f < n
invariant exists N . ~node.failed(N)

# 'after' a node has failed it cannot start working again
invariant node.failed_before_round(N) -> node.failed(N)

# 'after' a node has failed it will not send anything
invariant node.failed_before_round(N) -> ~node.incomming_message(M, N)

# If the decision round occured, a clean round must have occured
invariant system.decision_round -> system.clean_round_occured

\end{lstlisting}
\end{mdframed}

\section{Evaluation}

How viable is Ivy as a tool for algorithm designers to proof correctness
of their work? The goal when turning to a tool like Ivy is certainty of
correctness. But absolute certainty is never really achieved. When specifying the
algorithm in the Ivy language one can not trivially determine that the specification
is logically equivalent to the actual algorithm. This is compounded by the fact
that the Ivy language is structurally unlike classical procedural languages. Making it challenging to achieve a one-to-one translation from a classical distributed algorithm in a language like C. An example of this is the relatively compley quantifier-logic in assignments that update the global system state.  A similar thread is posed by $assume$ statements. The author of this paper originally modeled the line

\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=5, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
\begin{lstlisting}
node.w(N,V) := (node.w(N,V) | (~node.failed(N) & exists J . (node.incomming_message(N, J) & node.message(J, V))));
\end{lstlisting}
\end{mdframed}

as

\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=5, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
\begin{lstlisting}
 node.tmp_w(A,V) := node.w(A, V);
    node.w(A, V) := *;
    assume (node.tmp_w(A, V) | ((node.incomming_message(A, J) & node.message(J, V)))) <-> node.w(A, V);
\end{lstlisting}
\end{mdframed}
 and did not discover that the $assume$ is bogus and excludes important possibilities from being verified. The error only became apparent when it was discovered that correctness could be proven in the absence of an invariant that should have been absolutely necessary for induction.

- an invariant can exclude too much
--todo finde so eine invariante --

Furthermore, it has already been shown that the lack of easy iteration over types makes it difficult to achieve modularity. The lack of which introduces additional uncertainty about correct modelling, especially in distributed systems.

However, as has been pointed out, it is not just the algorithm that must be correctly specified, but the underlying framework is equally important to get right. In the presented specification it is possible to remove the \textit{simulate\_failure} action and still receive a positive check from the Ivy system. This is unsurprising, as we do not have any mechanism in place that checks that the system correctly models the possibility for node failure. E.g. to check that the simulate failure action allows a node to fail and send a message only to a subset of other nodes one can purposefully insert an assertion that only fails when the desired behaviour is possible, e.g. insert
\begin{mdframed}[backgroundcolor=light-gray, roundcorner=10pt,leftmargin=1, rightmargin=1, innerleftmargin=5, innertopmargin=15,innerbottommargin=15, outerlinewidth=1, linecolor=light-gray]
\begin{lstlisting}
ensure ~(node.failed(N) &
        (exists J . ~incoming_message(J, N) &
          exists K . incomming_message(K, N) & J~=N & K~=J))
\end{lstlisting}
\end{mdframed}
after the \textit{simulate\_failure} action. If Ivy does not find a counterexample to this assert, the specification is erroneous.  In light of this, it would be beneficial to have a built in mechanism to verify the possibility of something to occur. Allowing one to conveniently verify that the modeled system allows for the possibility of a node to fail and only send messages to a subset of processes as part of the final proof.

Of course a variety of bugs and exploits do not stem from a logically unsound algorithmic foundation but implementation errors. A wonderful example is the Heartbleed exploit in OpenSSL. It would have been no help to verify the simple Heartbleed protocol in Ivy before implementing it in C, since the error was in an implementation detail allowing a buffer underflow. The point is, the implementation of the algorithm must be tested through unit testing or bounded model checkers, despite successful verification in Ivy.

- induction is quite an unintuitive for beginners to get their head around

\section{Conclusion}
This paper gave a brief introduction into Controlled Natural Languages (CNLs)

%----------------------------------------------------------------------------------------
%	REFERENCE LIST
%----------------------------------------------------------------------------------------
\begin{thebibliography}{99}

    \bibitem{refLanguageDoc} http://microsoft.github.io/ivy/language.html
    \bibitem{refNancy} http://microsoft.github.io/ivy/language.html
\end{thebibliography}

%----------------------------------------------------------------------------------------

\end{document}